---
title: "Qualitative Activity Recognition of Weight Lifting Exercises"
author: "Nishanth"
output: html_document
---

## Summary
The goal is to determine the quality of the exercise (bicep curl) by monitoring various features like roll, pitch, yaw, accelerometer readings, gyroscope readings. Measurements are taken at elbow, arm and hip. The five classes are:  

1. **A** Exact bicep curl
2. **B** Throwing elbows to the front
3. **C** Lifting the dumbbell only halfway
4. **D** Lowering the dumbbell only halfway
5. **E** Throwing the hip forward

```{r,echo=FALSE}
setwd("F:/Workspace/Coursera/predmachlearn/")
```

## Loading the necessary packages and data
```{r,echo=TRUE,results='hide',message=F,warning=F,comment=NA}
library(caret)
library(data.table)
training = fread("pml-training.csv")
testing = fread("pml-testing.csv")
```


## Pre-processing
The training data has many missing values like `NA` and `""`. All the columns with missing values are removed from the training set. Remove the corresponding columns from test set too. First, remove `NA`s:  
```{r,echo=TRUE,message=F,warning=F,comment=NA}
na.cols = sapply(training, function(x) sum(is.na(x)))
tr = training[,which(na.cols == 0),with=FALSE]
ts = testing[,which(na.cols == 0),with=FALSE]
```

Then remove columns with empty strings `""`:  
```{r,echo=TRUE,message=F,warning=F,comment=NA}
na.cols = sapply(tr, function(x) sum((x=="")))
tr = tr[,which(na.cols == 0),with=FALSE]
ts = ts[,which(na.cols == 0),with=FALSE]
```

Next, remove columns like user name, id, time stamp which have no predictive power. Also convert the outcome variable to `factor` type:  
```{r,echo=TRUE,results='hide',message=F,warning=F,comment=NA}
tr[,`:=`(V1=NULL,user_name=NULL,cvtd_timestamp=NULL,
         classe=as.factor(classe),new_window=NULL)]
ts[,`:=`(V1=NULL,user_name=NULL,cvtd_timestamp=NULL,
         problem_id=NULL,new_window=NULL)]
```


## Predictive Modeling
For predictive modeling we use the random-forest algorithm. First divide the data into 70% training and 30% validation set. 
```{r,echo=TRUE,message=F,warning=F,comment=NA}
inTrain = createDataPartition(y=tr$classe,p=0.7,list=FALSE)[,1]
train = tr[inTrain,]
test = tr[-inTrain,]
```
Train the model using training set and predict using validation set. Check the out of sample error estimate using confusion matrix.
```{r,echo=TRUE,message=F,warning=F,comment=NA}
fit = train(classe~., data=train, method="rf", ntree=10)
pred = predict(fit, newdata=test)
confusionMatrix(test$classe, pred)
```

Finally, train the model using full training set. And generate the predictions for the original test set.  
```{r,eval=FALSE,echo=TRUE,message=F,warning=F,comment=NA}
final.fit = train(classe~., data=tr, method="rf", ntree=10)
predictions = predict(final.fit, newdata=testing)
```

